---
title: "Gun Violence Stats"
author: "Sean Nguyen"
date: "May 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(usmap)
library(knitr)
gv <- read_csv("gun-violence-data_01-2013_03-2018.csv")
```

## Incidents

### Total Incidents By State
```{r}
gv_by_state <- count(gv,state)
gv_by_state$state <- factor(gv_by_state$state, levels = gv_by_state$state[order(gv_by_state$n)])
gv_by_state_map <- gv_by_state %>% mutate(fips=vapply(gv_by_state$state, fips, character(1))) %>% select(fips,n)

plot_usmap(data=gv_by_state_map, regions="states", values="n") + scale_fill_continuous(low="yellow", high="red", name="Incidents") + theme(legend.position="right")

gv_by_state %>% arrange(desc(n)) %>% .[1:5,] %>% rename(Incidents=n, "Top 5 States"=state) %>% knitr::kable(.)
```

### Incidents Over Time
```{r}
ggplot(gv) + geom_freqpoly(mapping=aes(x=date), bins=50)
```

An initial look at the data shows relatively few incidents before 2014. This might be because the Gun Violence Archive, which is the source of this data set, was formed some time in 2013. Before the GVA's formation, the government and police might not have kept as many detailed records of gun violence.

```{r}
gv %>% filter(date < "2014-01-01") %>% ggplot + geom_freqpoly(mapping=aes(x=date))
```

### Most Violent Incidents
```{r}
gv %>% arrange(desc(n_killed)) %>% .[1:5,] %>% select(date, state, city_or_county, n_killed, n_injured) %>% rename(Date=date, State=state, City=city_or_county, Killed=n_killed, Injured=n_injured) %>% kable
```

While mass shootings grab headlines, only 27 out of the 260,000 reported incidents resulted in > 5 deaths (0.01%), and the majority of incidents cause zero deaths.

### People Killed Per Incident

```{r}
plot_colors <- c("green","red","red","red","red","red")
gv %>% filter(n_killed <= 5) %>% ggplot + geom_histogram(mapping=aes(x=n_killed),binwidth=1, fill=plot_colors) + geom_text(stat="count", mapping=aes(x=n_killed,label=..count..), vjust=-1) + scale_x_continuous(breaks=seq(from=0,to=5,by=1)) + scale_y_continuous(limits=(c(NA,250000))) + labs(x="# Killed", y="Occurrences") + scale_fill_manual(values=c("green","red","red","red","red","red"))
```

